---
title: "Data Scientists From Non-Computer Science Background"
author: "Muhsina Elampilakkattil"
date: "20 November 2018"
output: 
  html_document:
    toc: true
    fig_height: 5
    code_folding: hide
---
<hr/>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction 

"Data Scientist is ... "
Programming skills and statistical knowledge are the most essential skills of a data scientist.

"Can anyone from a non-computer science background be a data scientist?" "How can I be data scientist with a non-computer science background?" 
The most common questions from some aspiring data scientists. In fact, these were the same questions which I googled 7 months back. 
Data Scientist is the sexiest job of the 21st century.  There are so many people from diverse backgrounds trying to get this hottest job. Some of them think this job is meant for people from the computer science background. 
When I started thinking about a career in data science, I had lots of questions flying over my head.  What programming language do I have to learn? Which online source will help me learning to programme? From where I get public datasets to do projects? etc..
I thought it would be better if I get answers for these questions from people who are currently working as data scientists and from the non-computer science background. 
In this exploratory analysis, I will be trying to get answers for these questions from data scientists with non-computer science background. 

###About the Dataset 

 Kaggle conducted a worldwide survey to know about the state of data science and machine learning. The survey was live for one week in October. The datasets include 23,859 responses and 395 variables. These variables are the answers to survey questions. 

When survey respondents selected the "Other" category, an option was given for a text response. These text responses were separated and shuffled to protect user privacy.

We need to load multiple libraries and datasets to start the analysis.

```{r, message= FALSE, warning= FALSE}
library(tidyverse)
library(RColorBrewer)
library(ggthemes)
library(gridExtra)
library(cowplot)

#Load data
other<- read_csv("datasets/freeFormResponses.csv")
survey<- read_csv("datasets/multipleChoiceResponses.csv")
schema<- read_csv("datasets/SurveySchema.csv");

#Storing questions in a long data frame for reference
questions<- read_csv("datasets/multipleChoiceResponses.csv", n_max= 1)
questions<- gather(questions, column, descr)

#Remove the row with the long questions
survey<- survey[-1, ]
other<- other[-1, ]
```
###[Types of Data Scientists](https://blog.udacity.com/2018/01/4-types-data-science-jobs.html) 
Data science combines several disciplines, including statistics, data analysis, machine learning, and computer science. 
 Four types of data science jobs are,
** The Data Analyst
** The Data Engineer
** The Machine Learning Engineer
** The Data Science Generalist

Therefore, here we are grouping Data Scientists, Data Analysts, and Data Engineers together.  We are considering undergraduate major (Question 4) as the attribute to subset the data as respondents working as data scientists with a non-computer science major. We need to filter out responses with an option "I am a student" to the Question 7 (In what industry is your current employer/contract?) to ensure the respondent is a working as a data scientist. 

```{r}
#Data Scientists
survey%>%
filter(Q6 %in% c("Data Scientist", "Data Engineer", "Data Analyst"),
      !is.na(Q8),
       !is.na(Q5),
      Q7 != "I am a student") %>%  
  mutate(CS= ifelse(Q5=="Computer science (software engineering, etc.)", TRUE, FALSE)) %>%
  select(Q6, CS)   %>%
  group_by(Q6, CS)  %>%
summarize(n= n()) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
  ggplot(aes(x= reorder(Q6, n), y = n, fill= CS)) + 
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("dodgerblue4", "gray70")) +
  labs(title= "Data Scientists", x="", y="Count") +
        annotate("text", x = "Data Analyst", y = 1665, label = "73.8%", hjust= 2) +
        annotate("text", x = "Data Engineer", y = 667, label = "47.4%", hjust= 1.2) +
        annotate("text", x = "Data Scientist", y = 3739, label = "67.7%", hjust= 2) +
  coord_flip() + 
  theme_minimal(base_size= 10)

#Subset responses from non computer science major
NonCS<- survey%>%
filter(Q5 != "Computer science (software engineering, etc.)",
      Q6 %in% c("Data Scientist", "Data Engineer", "Data Analyst"),
      !is.na(Q8),
      Q7 != "I am a student")  




NonCS%>%
count(Q1) %>%
filter(!(Q1 %in% c("Prefer not to say", "Prefer to self-describe"))) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
  ggplot(aes(x= reorder(Q1, n),y=n)) +
  geom_bar(stat= "identity", fill= "blue") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Gender Distribution (Non- CS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#Country living in


NonCS%>%
count(Q3) %>%
filter(!(Q3 %in% c("Other", "I do not wish to disclose my location"))) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
head(6) %>%
  ggplot(aes(x= reorder(Q3, n), y= n)) +
  geom_bar(stat= "identity", fill= "blue") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Country (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)

#Employment Industry


NonCS%>%
count(Q7) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q7, n), y= n)) +
  geom_bar(stat= "identity", fill= "blue") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "EmploymentIndustry (Non CS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


NonCS%>%
count(Q4) %>%
filter(!(Q4 %in% c("Other", "I do not wish to disclose my location"))) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q4, n), y= n)) +
  geom_bar(stat= "identity", fill= "blue") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Country (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#Income


NonCS%>%
filter(!(Q9 %in% c(NA, "I do not wish to disclose my approximate yearly compensation"))) %>%
count(Q9) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q9, n), y= n)) +
  geom_bar(stat= "identity", fill= "blue") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Income (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#Activities important in the current role
(activities<- NonCS%>%
  select(matches("Q11_Part")) %>%
  gather(Part, Activities, na.rm=TRUE) %>%
  group_by(Activities) %>%
  summarize(count= n()) %>%
  arrange(desc(count)))


# Core wrapping function
wrap.it <- function(x, len)
{ 
  sapply(x, function(y) paste(strwrap(y, len), 
                              collapse = "\n"), 
         USE.NAMES = FALSE)
}


# Call this function with a list or vector
wrap.labels <- function(x, len)
{
  if (is.list(x))
  {
    lapply(x, wrap.it, len)
  } else {
    wrap.it(x, len)
  }
}


label <- wrap.labels(activities$Activities, 30)

activities%>%
mutate(pct = count/ sum(count) * 100 ) %>%
arrange(desc(count)) %>%
  ggplot(aes(x= reorder(Activities, desc(count)), y= count)) +
  geom_bar(stat= "identity", fill= "blue") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 3)+
  labs(title= "Activities important in the current role (NonCS)", x= "", y= "Count") +
    scale_x_discrete(labels= wrap.labels(activities$Activities, 30))+
  coord_flip() +
  theme_minimal(base_size= 8)


#Primary tool use at work
NonCS%>%
filter(!is.na(Q12_MULTIPLE_CHOICE)) %>%
count(Q12_MULTIPLE_CHOICE) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q12_MULTIPLE_CHOICE, n), y= n)) +
  geom_bar(stat= "identity", fill= "blue") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Primary tool use at work (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#Activities important in the current role
IDE<- NonCS%>%
  select(matches("Q13_Part")) %>%
  gather(Part, IDEs, na.rm=TRUE) %>%
  group_by(IDEs) %>%
  summarize(count= n()) %>%
  arrange(desc(count))

IDE%>%
mutate(pct = count/ sum(count) * 100 ) %>%
arrange(desc(count)) %>%
  ggplot(aes(x= reorder(IDEs, count), y= count)) +
  geom_bar(stat= "identity", fill= "burlywood1") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 3)+
  labs(title= "IDEs used at work (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)

#Notebooks used at work
notebooks<- NonCS%>%
  select(matches("Q14_Part")) %>%
  gather(Part, notebook, na.rm=TRUE) %>%
  group_by(notebook) %>%
  summarize(count= n()) %>%
  arrange(desc(count))

notebooks%>%
mutate(pct = count/ sum(count) * 100 ) %>%
arrange(desc(count)) %>%
  ggplot(aes(x= reorder(notebook, count), y= count)) +
  geom_bar(stat= "identity", fill= "burlywood1") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 3)+
  labs(title= "Notebooks used at work (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#Programming laguages used at work
lang<- NonCS%>%
  select(matches("Q16_Part")) %>%
  gather(Part, languages, na.rm=TRUE) %>%
  group_by(languages) %>%
  summarize(count= n()) %>%
  arrange(desc(count))

lang%>%
mutate(pct = count/ sum(count) * 100 ) %>%
arrange(desc(count)) %>%
  ggplot(aes(x= reorder(languages, count), y= count)) +
  geom_bar(stat= "identity", fill= "burlywood1") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 3)+
  labs(title= "Programming languages used at work (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


NonCS%>%
filter(!is.na(Q17)) %>%
count(Q17) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q17, n), y= n)) +
  geom_bar(stat= "identity", fill= "darksalmon") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Language use at work (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


NonCS%>%
filter(!is.na(Q18)) %>%
count(Q18) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q18, n), y= n)) +
  geom_bar(stat= "identity", fill= "darksalmon") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Language recommend (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#Machine learning framework
mlFramework<- NonCS%>%
  select(matches("Q19_Part")) %>%
  gather(Part, framework, na.rm=TRUE) %>%
  group_by(framework) %>%
  summarize(count= n()) %>%
  arrange(desc(count))

mlFramework%>%
mutate(pct = count/ sum(count) * 100 ) %>%
arrange(desc(count)) %>%
  ggplot(aes(x= reorder(framework, count), y= count)) +
  geom_bar(stat= "identity", fill= "burlywood1") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 3)+
  labs(title= "ML Framework used at work (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


NonCS%>%
filter(!is.na(Q20)) %>%
count(Q20) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q20, n), y= n)) +
  geom_bar(stat= "identity", fill= "darksalmon") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "ML framework mostly used (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


dataVizLib<- NonCS%>%
  select(matches("Q21_Part")) %>%
  gather(Part, library, na.rm=TRUE) %>%
  group_by(library) %>%
  summarize(count= n()) %>%
  arrange(desc(count))

dataVizLib%>%
mutate(pct = count/ sum(count) * 100 ) %>%
arrange(desc(count)) %>%
  ggplot(aes(x= reorder(library, count), y= count)) +
  geom_bar(stat= "identity", fill= "burlywood1") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 3)+
  labs(title= "Dataviz libraries used at work (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


NonCS%>%
filter(!is.na(Q22)) %>%
count(Q22) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q22, n), y= n)) +
  geom_bar(stat= "identity", fill= "darksalmon") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "DataViz library mostly used (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


NonCS%>%
filter(!is.na(Q22)) %>%
count(Q22) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q22, n), y= n)) +
  geom_bar(stat= "identity", fill= "darksalmon") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "DataViz library mostly used (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


NonCS%>%
filter(!is.na(Q26)) %>%
count(Q26) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q26, n), y= n)) +
  geom_bar(stat= "identity", fill= "darksalmon") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Are you a DS? (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#Data used
data<- NonCS%>%
  select(matches("Q31_Part")) %>%
  gather(Part, type, na.rm=TRUE) %>%
  group_by(type) %>%
  summarize(count= n()) %>%
  arrange(desc(count))

data%>%
mutate(pct = count/ sum(count) * 100 ) %>%
arrange(desc(count)) %>%
  ggplot(aes(x= reorder(type, count), y= count)) +
  geom_bar(stat= "identity", fill= "burlywood1") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 3)+
  labs(title= "data types used at work (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


NonCS%>%
filter(!is.na(Q32)) %>%
count(Q32) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q32, n), y= n)) +
  geom_bar(stat= "identity", fill= "darksalmon") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Data types mostly deal with (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#Datasets
datasets<- NonCS%>%
  select(matches("Q33_Part")) %>%
  gather(Part, datasets, na.rm=TRUE) %>%
  group_by(datasets) %>%
  summarize(count= n()) %>%
  arrange(desc(count))

datasets%>%
mutate(pct = count/ sum(count) * 100 ) %>%
arrange(desc(count)) %>%
  ggplot(aes(x= reorder(datasets, count), y= count)) +
  geom_bar(stat= "identity", fill= "burlywood1") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 3)+
  labs(title= "public datasets (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#online platform
online<- NonCS%>%
  select(matches("Q36_Part")) %>%
  gather(Part, platform, na.rm=TRUE) %>%
  group_by(platform) %>%
  summarize(count= n()) %>%
  arrange(desc(count))

online%>%
mutate(pct = count/ sum(count) * 100 ) %>%
arrange(desc(count)) %>%
  ggplot(aes(x= reorder(platform, count), y= count)) +
  geom_bar(stat= "identity", fill= "burlywood1") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 3)+
  labs(title= "online platforms (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


NonCS%>%
filter(!is.na(Q37)) %>%
count(Q37) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q37, n), y= n)) +
  geom_bar(stat= "identity", fill= "darksalmon") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Online platform mostly used (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)

#media
media<- NonCS%>%
  select(matches("Q38_Part")) %>%
  gather(Part, media, na.rm=TRUE) %>%
  group_by(media) %>%
  summarize(count= n()) %>%
  arrange(desc(count))

media%>%
mutate(pct = count/ sum(count) * 100 ) %>%
arrange(desc(count)) %>%
  ggplot(aes(x= reorder(media, count), y= count)) +
  geom_bar(stat= "identity", fill= "burlywood1") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 3)+
  labs(title= "media source (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)



