---
title: "kaggle_survey_2018"
author: "Muhsina Elampilakkattil"
date: "20 November 2018"
output: 
  html_document:
    toc: true
    fig_height: 5
    code_folding: hide
---
<hr/>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction



#Load Libraries


```{r, message= FALSE, warning= FALSE}
library(tidyverse)
library(RColorBrewer)
library(ggthemes)
library(gridExtra)
library(cowplot)

#Load data
other<- read_csv("datasets/freeFormResponses.csv")
survey<- read_csv("datasets/multipleChoiceResponses.csv")
schema<- read_csv("datasets/SurveySchema.csv");

#Storing questions in a long data frame for reference
questions<- read_csv("datasets/multipleChoiceResponses.csv", n_max= 1)
questions<- gather(questions, column, descr)



survey<- survey[-1, ]
other<- other[-1, ]

survey%>%
count(Q1) %>%
filter(!(Q1 %in% c("Prefer not to say", "Prefer to self-describe"))) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
  ggplot(aes(x= reorder(Q1, n),y=n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Gender Distribution", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)

survey%>%
count(Q3) %>%
filter(!(Q3 %in% c("Other", "I do not wish to disclose my location"))) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
head(6) %>%
  ggplot(aes(x= reorder(Q3, n), y= n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Country", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#School major
survey%>%
count(Q5) %>%
filter(!(Q5 %in% c("Other", "I never declared a major")),
      !is.na(Q5)) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
  ggplot(aes(x= reorder(Q5, n),y=n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "School Major", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)

#JobRole
survey%>%
count(Q6) %>%
filter(!is.na(Q6)) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
  ggplot(aes(x= reorder(Q6, n),y=n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Current Job title/ Role", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)

CS<-  survey%>%
filter(Q5 == "Computer science (software engineering, etc.)",
      Q6 %in% c("Data Scientist", "Data Engineer", "Data Analyst"),
      !is.na(Q8),
      Q7 != "I am a student")  


NonCS<- survey%>%
filter(Q5 != "Computer science (software engineering, etc.)",
      Q6 %in% c("Data Scientist", "Data Engineer", "Data Analyst"),
      !is.na(Q8),
      Q7 != "I am a student")  


CS%>%
count(Q1) %>%
filter(!(Q1 %in% c("Prefer not to say", "Prefer to self-describe"))) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
  ggplot(aes(x= reorder(Q1, n),y=n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Gender Distribution (CS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)

NonCS%>%
count(Q1) %>%
filter(!(Q1 %in% c("Prefer not to say", "Prefer to self-describe"))) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
  ggplot(aes(x= reorder(Q1, n),y=n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Gender Distribution (Non- CS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#Country living in
CS%>%
count(Q3) %>%
filter(!(Q3 %in% c("Other", "I do not wish to disclose my location"))) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
head(6) %>%
  ggplot(aes(x= reorder(Q3, n), y= n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Country (CS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)

NonCS%>%
count(Q3) %>%
filter(!(Q3 %in% c("Other", "I do not wish to disclose my location"))) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
head(6) %>%
  ggplot(aes(x= reorder(Q3, n), y= n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Country (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)

#Employment Industry
CS%>%
count(Q7) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q7, n), y= n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "EmploymentIndustry (CS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)

NonCS%>%
count(Q7) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q7, n), y= n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "EmploymentIndustry (Non CS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


NonCS%>%
count(Q4) %>%
filter(!(Q4 %in% c("Other", "I do not wish to disclose my location"))) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q4, n), y= n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Country (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#Income
CS%>%
filter(!(Q9 %in% c(NA, "I do not wish to disclose my approximate yearly compensation"))) %>%
count(Q9) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q9, n), y= n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Income (CS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)

NonCS%>%
filter(!(Q9 %in% c(NA, "I do not wish to disclose my approximate yearly compensation"))) %>%
count(Q9) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q9, n), y= n)) +
  geom_bar(stat= "identity", fill= fillBlue) +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Income (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)


#Activities important in the current role
(activities<- NonCS%>%
  select(matches("Q11_Part")) %>%
  gather(Part, Activities, na.rm=TRUE) %>%
  group_by(Activities) %>%
  summarize(count= n()) %>%
  arrange(desc(count)))


# Core wrapping function
wrap.it <- function(x, len)
{ 
  sapply(x, function(y) paste(strwrap(y, len), 
                              collapse = "\n"), 
         USE.NAMES = FALSE)
}


# Call this function with a list or vector
wrap.labels <- function(x, len)
{
  if (is.list(x))
  {
    lapply(x, wrap.it, len)
  } else {
    wrap.it(x, len)
  }
}


label <- wrap.labels(activities$Activities, 30)

activities%>%
mutate(pct = count/ sum(count) * 100 ) %>%
arrange(desc(count)) %>%
  ggplot(aes(x= reorder(Activities, desc(count)), y= count)) +
  geom_bar(stat= "identity", fill= "blue") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 3)+
  labs(title= "Activities important in the current role (NonCS)", x= "", y= "Count") +
    scale_x_discrete(labels= wrap.labels(activities$Activities, 30))+
  coord_flip() +
  theme_minimal(base_size= 8)


#Primary tool use at work
NonCS%>%
filter(!is.na(Q12_MULTIPLE_CHOICE))
count(Q12_MULTIPLE_CHOICE) %>%
mutate(pct = n/ sum(n) * 100 ) %>%
arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Q12_MULTIPLE_CHOICE, n), y= n)) +
  geom_bar(stat= "identity", fill= "blue") +
  geom_text(aes(label= paste0(round(pct,2),"%", sep="")),
      position= position_dodge(0), color= "black", size= 2.5)+
  labs(title= "Primary tool use at work (NonCS)", x= "", y= "Count") +
  coord_flip() +
  theme_minimal(base_size= 8)

